JWST Noise Floor I: Systematic Error Sources in JWST NIRCam Time Series

### 2020-05-28 ms.tex
incorporated ME comments

Given the finite lifetime of the JWST mission set by onboard fuel, it is important to understand the systematic errors before launch to maximize the science return of the observatory.
We discuss the major sources of systematic errors that can affect time-series observations with the NIRCam instrument.
Parallel efforts to characterize other instrument's noise sources are being done, such as with MIRI's detectors \citep{matuso2019siAsDetectorStability}.

@@

The following sources of noise introduce random variations to time series:
\begin{enumerate}[noitemsep]
	\item \textbf{Pre-amplifier reset offsets:} (a SIDECAR ASIC-related effect) \label{it:preAmpOffsets}
The resets cause discontinuities in the pedestal level of all pixels between frames.
Pre-amp resets can also appear in slope images (DN/second) because NIRCam's amplifiers are reset once per frame (not once per integration).
This is an effect where averaging over more pixels does not decrease the noise, but reference pixel or background subtraction can efficiently remove the pre-amp offsets.
These are discussed further in Section \ref{sec:preAmp}.
	\item \textbf{Amplifier Boundary Discontinuities} The NIRCam detectors can be read out in full frame, stripe and window modes. The full frame and stripe modes make use of 4 amplifiers that can record or reset the accumulated electrons in 4 pixels simultaneously in parallel. The parallel operation of the amplifiers reduces the frame time and reset time by a factor of 4 but can introduce subtle voltage biases between regions of the detector. If these voltage biases are not corrected, they can potentially introduce discontinuities in a spectrum, which crosses amplifier boundaries.
	The amplifier boundary discontinuities are a consequence of the pre-amplifier reset offsets above.
	\item \textbf{1/f noise:} The JWST readout integrated circuits (ROICs) have a p-type metal-oxide semiconductor field-effect transistor (PFET) source follower that introduces noise to the images as well as DC biases in the ASIC electronics \citep{rauscher2011irsSquared}.
	This is a noise source where most of the noise power is concentrated at low frequencies.
	1/f noise causes spatial correlations primarily in the fast-read direction (along detector rows). It is highly correlated between amplifiers, suggesting that it is caused by a common reference voltage. 1/f noise can be reduced by subtracting the values from background pixels or reference pixels.
	\item \textbf{Even/odd offsets:} Alternating columns have different bias levels. Additionally, there are offsets between columns that can change from frame to frame, especially on NIRCam's LW detector.
	These can be reduced with reference pixels or background subtraction.
	\item \textbf{kTC noise:} Each reset at the beginning of an integration introduces thermal noise associated with the unknown amount of charge stored in a pixel at the reset voltage level.
	Fortunately, the bias offset produced by kTC noise will easily subtract out when fitting a slope image from the multiple samples up the ramp.
	This noise would  only become relevant when attempting to make use of the first group of a ramp (such as when only 1 group is commanded).
	\item \textbf{Elevated Columns} There is correlated noise along a column in many frames. These elevated columns can sometimes even move with time.
	\item \textbf{Random Telegraph Noise (RTN)} Some of the pixels will exhibit spontaneous changes jumps in signal even with no illumination. Fortunately, RTN appears in only $\sim$1000 pixels out of 4$\times 10^6$ on the array and exhibited 1 jump in a 20 minute exposure.
\end{enumerate}

In addition to these random effects, there are also systematic effects that are time-variable only due variations in other variables such as temperature and telescope pointing:
\begin{enumerate}[resume]
	\item \textbf{Intrapixel sensitivity:} Telescope pointing drifts and jitter can impact the amount of flux measured if the response within pixels is not perfectly uniform.
	This has a particularly strong impact on the time series from the Spitzer IRAC instrument \citep{ingalls2016spitzerRepeatability}.
	It can also appear on JWST detectors because the sub-pixel flat field is not uniform.
	Crosshatching patterns on HgCdTe detectors (which all JWST near-infrared instruments use) exist at the sub-pixel level  \citep{shapiro2018crosshatch,ninan2019crosshatchHPF}
	\item \textbf{Detector Temperature Fluctuations:} The JWST detectors are sensitive to temperature changes on the focal plane arrays. For example, laboratory tests show that 100 mK temperature fluctuations can result in $\sim$ 80 DN changes on the detector that are not corrected by reference pixels \citep{hall2005jwstArrays}. Fortunately, the detectors are actively thermally controlled to keep the temperature fluctuations to mostly $\lesssim$ 1 mK.
The one exception is that the Long Wavelength A-side detector (A5) exhibited $\sim$ 20 min oscillations with an amplitude of 15-20 mK when using the Temperature Monitor Control (TMC) 1 active thermal control.
Fortunately, the TMC 2 control does not undergo these oscillations and will be employed in orbit.
If necessary, the effect of temperature fluctuations can be calibrated to $\pm 1 e^-$ for excursions less than $\pm$ 50 mK \citep{hall2005jwstArrays}.
	\item \textbf{SIDECAR ASIC Temperature Fluctuations:} The SIDECAR ASIC temperature will change with the operating mode. For example, in the CV3 stability test, the temperature of the A SIDECAR ASIC changed by $\sim$80 mK when the readout mode was switched from full frame to subarray.
	This temperature change is expected to be smaller in flight due to adjustments in the operations to run all four amplifiers regardless of how many are used to collect science data.
	\item \textbf{Detector Fringing} The thin $\sim 6\mu$m layer of the H2RG detectors or other optics can cause interference patterns on the array due to constructive and destructive interference of light at wavelengths similar to the thickness of the layers. Older HgCdTe detectors contained a CdZnTe layer that caused a prominent fringing pattern \citep{finger2008HgCdTeCharacterization}, but JWST detectors have the CdZnTe removed so the fringe pattern comes from a different optical layer. If the detector fringe pattern changes with time, it will change the throughput of a grism image.
	\item \textbf{Charge Trapping} A major correlated error source on HST WFC3 is the ``hook effect'' \citep{berta2012flat_gj1214}.
	The hook effects is due to charge can be trapped within the detector's depletion region (initially lowering the flux) and then released at later times (increasing the flux) \citep{zhou2017chargeTrap}.
	JWST's NIRCam instrument has very lower rates of persistence ($<$1 DN/sec) at 100 seconds after saturation on all detectors \citep{leisenring2016persistence}, so it is expected that charge trapping will be a minor concern for JWST's HgCdTe detectors.
\end{enumerate}

@@

\paragraph{Reference Channel Subtraction}
We note that the 4 channels can sometimes exhibit the same 1/f correlation behavior, as shown in Figure \ref{fig:darkPixelTimeSeries}.
Another option to remove 1/f noise is to subtract the illuminated amplifier with a target in it by un-illuminated channels.
We subtract the first amplifier's time series (pixels X=1 through X=512) from all amplifiers.
Of course, this means that this amplifier is perfectly subtracted and no noise or signal remains in this 1/4 of the image.
The subtraction step should increase the read noise on an individual pixel by $\sim \sqrt{2}$ because the errors from the subtracted pixel should add in quadrature to the original pixels.
This increase in read noise on a per-pixel level might be acceptable if it decreases the correlated errors along the rows.
On the other hand, if the each amplifier exhibits its own individual correlated noise behavior, the reference amplifier will fail to subtract out all correlated noise and may introduce more noise.

>>>>

Given the finite lifetime of the JWST mission set by onboard fuel, it is important to characterize the systematic errors before launch to maximize the science return of the observatory.
We discuss the known sources of systematic errors that can affect time-series observations with the NIRCam instrument.
Parallel efforts to characterize other instrument's noise sources are being done, such as with MIRI's detectors \citep{matuso2019siAsDetectorStability}.

@@

The following sources of noise introduce random variations to time series:
\begin{enumerate}[noitemsep]
	\item \textbf{Pre-amplifier reset offsets:} (a SIDECAR ASIC-related effect) \label{it:preAmpOffsets}
The resets cause discontinuities in the pedestal level of all pixels between frames.
Pre-amp resets can also appear in slope images (DN/second) because NIRCam's amplifiers are reset once per frame (not once per integration).
This is an effect where averaging over more pixels does not decrease the noise, but reference pixel or background subtraction can efficiently remove the pre-amp offsets.
These are discussed further in Section \ref{sec:preAmp}.
	\item \textbf{Amplifier Boundary Discontinuities} The NIRCam detectors can be read out in full frame, stripe and window modes. The full frame and stripe modes make use of 4 amplifiers that can record or reset the accumulated electrons in 4 pixels simultaneously in parallel. The parallel operation of the amplifiers reduces the frame time and reset time by a factor of 4 but can introduce subtle voltage biases between readout amplifiers of the detector. If these voltage biases are not corrected, they can potentially introduce discontinuities in a spectrum, which crosses amplifier boundaries.
	The amplifier boundary discontinuities are a consequence of the pre-amplifier reset offsets above.
	\item \textbf{1/f noise:} The JWST readout integrated circuits (ROICs) have a p-type metal-oxide semiconductor field-effect transistor (PFET) source follower that introduces noise to the images as well as DC biases in the ASIC electronics \citep{rauscher2011irsSquared}.
	This is a noise source where most of the noise power is concentrated at low frequencies.
	1/f noise causes spatial correlations primarily in the fast-read direction (along detector rows). It is highly correlated between amplifiers, suggesting that it is caused by a common reference voltage. 1/f noise can be reduced by subtracting the values from background pixels or reference pixels.
	\item \textbf{Even/odd offsets:} Alternating columns have different bias levels. Additionally, there are offsets between columns that can change from frame to frame, especially on NIRCam's LW detector.
	These can be reduced with reference pixels or background subtraction.
	\item \textbf{kTC noise:} Each reset at the beginning of an integration introduces thermal noise associated with the unknown amount of charge stored in a pixel at the reset voltage level.
	Fortunately, the bias offset produced by kTC noise will easily subtract out when fitting a slope image from the multiple samples up the ramp.
	This noise would  only become relevant when attempting to make use of the first group of a ramp (such as when only 1 group is commanded).
	\item \textbf{Elevated Columns} There is correlated noise along a column in many frames. These elevated columns can sometimes even move with time.
	\item \textbf{Random Telegraph Noise (RTN)} Some of the pixels will exhibit spontaneous changes jumps in signal even with no illumination. Fortunately, RTN appears in only $\sim$1000 pixels out of 4$\times 10^6$ on the array and exhibited 1 jump in a 20 minute exposure.
\end{enumerate}

In addition to these random effects, there are also systematic effects that are time-variable only due variations in other variables such as temperature and telescope pointing:
\begin{enumerate}[resume]
	\item \textbf{Intrapixel sensitivity:} Telescope pointing drifts and jitter can impact the amount of flux measured if the response within pixels is not perfectly uniform.
	This has a particularly strong impact on the time series from the Spitzer IRAC instrument \citep{ingalls2016spitzerRepeatability}.
	It can also appear on JWST detectors because the sub-pixel flat field is not uniform.
	Crosshatching patterns on HgCdTe detectors (which all JWST near-infrared instruments use) exist at the sub-pixel level  \citep{shapiro2018crosshatch,ninan2019crosshatchHPF}
	\item \textbf{Detector Temperature Fluctuations:} The JWST detectors are sensitive to temperature changes on the focal plane arrays. For example, laboratory tests show that 100 mK temperature fluctuations can result in $\sim$ 80 DN changes on the detector that are not corrected by reference pixels \citep{hall2005jwstArrays}. Fortunately, the detectors are actively thermally controlled to keep the temperature fluctuations to mostly $\lesssim$ 1 mK.
The one exception is that the Long Wavelength A-side detector (A5) exhibited $\sim$ 20 min oscillations with an amplitude of 15-20 mK when using the Temperature Monitor Control (TMC) 1 active thermal control.
Fortunately, the TMC 2 control (with an independent set of hardware) does not undergo these oscillations and will be employed in orbit.
If necessary, the effect of temperature fluctuations can be calibrated to $\pm 1 e^-$ for excursions less than $\pm$ 50 mK \citep{hall2005jwstArrays}.
	\item \textbf{SIDECAR ASIC Temperature Fluctuations:} The SIDECAR ASIC temperature will change with the operating mode. For example, in the CV3 stability test, the temperature of the A SIDECAR ASIC changed by $\sim$80 mK when the readout mode was switched from full frame to subarray.
	This temperature change is expected to be smaller in flight due to adjustments in the operations to run all four amplifiers regardless of how many are used to collect science data.
	\item \textbf{Detector Fringing} The thin $\sim 6\mu$m layer of the H2RG detectors or other optics can cause interference patterns on the array due to constructive and destructive interference of light at wavelengths similar to the thickness of the layers. Older HgCdTe detectors contained a CdZnTe layer that caused a prominent fringing pattern \citep{finger2008HgCdTeCharacterization}, but JWST detectors have the CdZnTe removed so the fringe pattern comes from a different optical layer. If the detector fringe pattern changes with time, it will change the throughput of a grism image.
	\item \textbf{Charge Trapping} A major correlated error source on HST WFC3 is the ``hook effect'' \citep{berta2012flat_gj1214}.
	The hook effects is due to charge can be trapped within the detector's depletion region (initially lowering the flux) and then released at later times (increasing the flux) \citep{zhou2017chargeTrap}.
	JWST's NIRCam instrument has very lower rates of persistence ($<$1 DN/sec) at 100 seconds after saturation on all detectors \citep{leisenring2016persistence}, so it is expected that charge trapping will be a minor concern for JWST's HgCdTe detectors.
	\item \textbf{Reciprocity Faillure:} The response of a HgCdTe detector can change with the brightness of a source, which is called reciprocity failure \citep{biesiadzinski2011reciprocityFailure}.
	The brightness of a star and planet system does not change significantly with time, so it will not significantly affect the {\it precision} of a lightcurve, but could affect {\it accuracy} because the filter bandpass and stellar spectral energy distribution.
	It is possible that reciprocity failure could introduce a systematic offset, slope or curvature that depends on the count rate of the detector.
\end{enumerate}

@@

\paragraph{Reference Channel Subtraction}
We note that the 4 channels can sometimes exhibit the same 1/f correlation behavior, as shown in Figure \ref{fig:darkPixelTimeSeries}.
Another option to remove 1/f noise is to subtract the 3 out of 4 output amplifiers that have the source on them by the 1 output amplifier without the target.
We subtract the first amplifier's time series (pixels X=1 through X=512) from all amplifiers by shifting and flipping the 2048$\times$256 image sub-section to the corners of the 4 amplifiers at (X=1, X=513, X=1025 and X=1537).
Of course, this means that this amplifier is perfectly subtracted and no noise or signal remains in this 1/4 of the image.
The subtraction step should increase the read noise on an individual pixel (for the remaining 3 output amplifiers) by $\sim \sqrt{2}$ because the errors from the subtracted pixel should add in quadrature to the original pixels.
This increase in read noise on a per-pixel level might be acceptable if it decreases the correlated errors along the rows.
On the other hand, if the each amplifier exhibits its own individual correlated noise behavior, the reference amplifier will fail to subtract out all correlated noise and may introduce more noise.
